Issues with Dynamical Measurements Across the Disk

### 2019-02-25 paper/ms.tex
changed simulation names to sans serif

\section{Azimuthal midplane variations at the Solar Circle: estimates from
FIRE} \label{sec:local_fire}
\subsection{Description of Simulation} \label{ssec:cosmozoom}
Here we briefly describe the cosmological-hydrodynamical zoom-in simulations
used in this work. Cosmological zoom-ins allow one to simulate a selected
region at high resolution embedded in a low-resolution cosmological background
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added
a hydrodynamical simulation of the baryonic component with recipes for star
formation and feedback. The FIRE collaboration has simulated a number of Milky
Way-mass galaxies using this technique as part of the {\em Latte} suite of
FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. For this
work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S},
which show broad agreement of many of their global properties with
observations of the Milky Way. The $\z=0$ snapshots\footnote{In this work, to
avoid confusion with the vertical height $z$, we refer to cosmological
redshift as $\z$.} of these three simulations, called m12i, m12f, and m12m for
shorthand, are publicly available alongside associated mock {\em Gaia} DR2
catalogues generated from them.\footnote{available at
\url{http://ananke.hub.yt}}

These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
gas particles of mass $\sim 7000 \text{--} 20,000\,\Msun$, and star particles
of mass $\sim 5000 \text{--} 7000\,
\Msun$,\footnote{In fact, the simulation m12i contained a gas
splitting bug which causes a small number of gas and star particles to have
larger than typical masses. These constitute only $\sim0.2\%$ of the star
particles in the $\z=0$ snapshots, and are only a factor of a few more massive
than typical star particles. Since our results are broadly consistent between
the three simulations, we ignore this minor complication.} with the lower end
coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
respectively.\footnote{This is $2.8$ times the often-quoted
Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
median softening length for gas particles around roughly solar positions (with
cylindrical radii within $500\,\pc$ of $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is
$98\,\pc$. %***see Issue***

The softening lengths used in the simulations can affect the ability to
resolve the very thinnest planar structures, which in turn can affect how much
the density-based midplane varies as a function of azimuth. The Milky Way's
dense, star-forming gas disk is thought to have a scale height of about 40 pc (XXX times the
median gas softening length; \citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of about 300 pc (30 times the stellar softening length; \citealt{2008ApJ...673..864J}). We therefore expect that
resolution effects may still be affecting the scale heights of these
components in the simulations, and indeed the gas and stellar scale heights of
the simulated galaxies are equal to or larger than the MW's. However, in this
work we are considering azimuthal variations in the midplane \emph{as defined
by the mass density}, which will subsequently be lower in the simulations than
in the MW: within 200 pc of the Sun the measured gas and stellar volume
densities in the MW are roughly comparable at $\sim .04\ \Msun\
\unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
Table 3 of \citealt{2018arXiv180610564S}). We thus expect that the degree of
midplane variation we find in the simulated systems is likely to be
\emph{less} than in the real MW, since on the smallest scales the simulations
are \emph{smoother} than the real Galaxy.

Cosmological simulations of Milky Way-mass galaxies are not perfect
representations of the true Milky Way in other ways as well, as discussed in
\citet{2018arXiv180610564S}. For instance, the velocity structure of m12i is
closer to M31's than the Milky Way's (Loebman et al. in prep). However, in
this work we are most interested in the global properties of the potential,
and specifically in deviations from axisymmetry. From this perspective, the
simulated galaxies are actually \emph{more} axisymmetric than we might expect
of the Milky Way. While they have prominent spiral arms, none has as strong a
bar as the MW does at present day, and none has a nearby companion like the
Large Magellanic Cloud. One of the three we consider (\mf) does have an
ongoing interaction with a satellite galaxy similar to Sagittarius, which has
punched through the galactic disk outside the solar circle and induced some
warping. However, we expect that, as with the use of finite smoothing lengths,
the degree of midplane variation in the simulations is likely a factor of a
few lower than what could be expected in the MW.

@@

To allow for potential small inaccuracies in the determination of the original fiducial coordinate system,
we also subtract the best fit $\text{A} \sin{\left(\phi + \text{B}\right)} +
\text{C}$ curve from the midplane as a function of azimuth. For $\text{A}$ the
values are $-174$, $62$, and $3.8\,\pc$, for $\text{B}$ the values are $0.3$,
$0.65$, and $0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$,
and $-15\,\pc$ for m12i, m12f, and m12m, respectively. For the assumed solar
radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
from the values of $\text{A}$ --- we compute $1.22$, $0.43$, and $0.03\,\deg$
for m12i, m12f, and m12m. These angle offsets are consistent with the values
given in \citet{2018arXiv180610564S} for the difference between the $z$-axis
as defined by the gas and stars.

Figure~\ref{fig:midplane} shows the inferred midplane our imaginary observer
would make as a function of azimuth for each galaxy. The 5--95 percentile range from the bootstrap procedure is shown as the dashed-line error bars. The $90\%$ interquartile
range for each galaxy is $185$, $162$, $84\,\pc$ for m12i, m12f, and m12m.

\begin{figure*}
\plotone{fig/midplane_fit.pdf}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for the three FIRE galaxies m12i, m21f, and m12m ({\em left},
{\em center}, and {\em right} panels). The local midplane is determined at a
position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
using the new height $10$ times to converge on the local midplane height. In
order to allow for the possibility that the fiducial Galactocentric coordinate
system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
each panel --- this figure is reproduced with the original midplane
determination (i.e. before subtracting the best fit sine curve) in
Appendix~\Gus{x}. We then bootstrap resample $1000$ times on all stars within
a $2\,\kpc$ height of the fiducial midplane to determine error bars (95\uth
and 5\uth percentiles), which we report as dashed lines.}
\label{fig:midplane}
\end{figure*}

@@

\subsection{Applicability to the Milky Way} \label{ssec:is_it_real}
As we briefly discussed in Section~\ref{ssec:cosmozoom}, cosmological zoom-in
simulations of Milky Way-mass galaxies are not perfect representations of the
Milky Way, and do not reproduce every available observation of the Milky Way.
For instance, the velocity dispersion-age relation in m12i, m12f, and m12m is
closer to that of M31 \citep{2018arXiv180610564S}.

In order to assess the applicability of our work to the real Milky Way, we
consider a subset of Table~3 from \citet{2018arXiv180610564S} --- our
Table~\ref{tab:scale_height} shows the stellar scale heights of m12i, m12f,
and m12m and the current best estimates for the Milky Way. One noticeable
feature is that the FIRE galaxies are more ``puffy'' than the Milky Way.
\Gus{Maybe FIRE collaboration has some insight into the mechanism for this?}.
This would seem to suggest that the midplane estimates from
Figure~\ref{fig:midplane} over-estimate the true variation in the Milky Way.

\begin{deluxetable}{ccc}
\tablecaption{Stellar disk scale heights of the real Milky Way and the
simulated galaxies considered in this work.\label{tab:scale_height}}
\tablehead{\colhead{galaxy} & \colhead{thin disk} & \colhead{thick disk} \\ 
\colhead{} & \colhead{(pc)} & \colhead{(pc)} } 
\startdata
MW\tablenotemark{a} & 300 & 900 \\
m12i\tablenotemark{b} & 480 & 2000 \\
m12f\tablenotemark{b} & 440 & 1280 \\
m12m\tablenotemark{b} & 290 & 1030 \\
\enddata
% \tablerefs{\citet{2016ARAA..54..529B,2018arXiv180610564S,2008ApJ...673..864J}}

@@

Another important aspect to consider when comparing the simulated galaxies to
the Milky Way is the environment/merger history. For instance, m12i and m12m
both lack a massive neighbor analogous to the LMC. On the other hand, m12f has
had a recenter merger with \Gus{something}. \Gus{Expand more here - how to
incorporate merger trees?}

@@

Figure~\ref{fig:Rn_mc} is most informative for the thin orbit (\thincolor).
The neighborhood around $\sim100\Msun$ clusters can be as small as
$\sim300\,\pc$ but up to $\sim1.2\,\kpc$ for m12i and m12f, though the
situation is marginally better for m12m with values ranging from
$\sim0.8\,\kpc$ to $\sim3\,\kpc$. The neighborhood values steadily increase
with $m_c$, topping out at $\sim5\,\kpc$ for $\sim10^4\,\Msun$ clusters in
m12i and m12f. In m12m, the neighborhood value reaches the maximum allowed of
$16.4\,\kpc$ for clusters of mass $\sim800\,\Msun$. The interpretation of this
is that the expected action space dispersion is so high for such high mass
clusters that the measured $z$~offset is not large enough throughout the
entire galaxy to be important.

For the thick-disk orbit (brown), the neighborhood is at the maximum value for
nearly all masses, except at the very low-mass end in m12i.

We repeat this entire procedure for each of the orbits of the real open
clusters in Table~\ref{tab:real_clusters} using the midplane variations from
m12i. We use m12i as a conservative choice, since it provides the greatest
midplane variation (Figure~\ref{fig:midplane}). We report the neighborhood
values for the mass of each cluster in the neighborhood column of
Table~\ref{tab:real_clusters}. Except for alphaPer, we find that the
neigbhorhood around each cluster is typically $\gtrsim 500\,\pc$. This
analysis assumes that the midplane from m12i can be accurately applied to
Milky Way orbits, which is not necessarily true. However, our analysis is
meant to be more illustrative than quantitative and these neighborhood numbers
should be understood as such.

\begin{figure*}
\plotone{fig/Rn_vs_mc.pdf}
\caption{The neighborhood around an open cluster of mass $m_c$ for the thin,
and thick orbits (\thincolor and \thickcolor, respectively; see
Table~\ref{tab:orbits}) and for each of the FIRE galaxies considered here ---
m12i, m12f, and m12m ({\em left}, {\em center}, and {\em right} panels,
respectively). Given the $z$ offset for a cluster of mass $m_c$ computed in
Figure~\ref{fig:cluster_offset}, we convert this to a chord length using the
result from Figure~\ref{fig:range_deltaphi}. We interpret this chord length as
the ``neighborhood'' around the cluster, i.e. the distance from the cluster
one could go before the $J_z$ error induced by our midplane effect is
comparable to the intrinsic action dispersion of the cluster. When the $z$
offset is larger than the maximum range in Figure~\ref{fig:range_deltaphi}, we
report the maximum chord length ($2\times8.2\,\kpc$). We aso report the same
procedure but using the $\pm1\sigma$ lines from
Figure~\ref{fig:range_deltaphi}, which we plot here as dashed lines. For the
thick orbit (\thickcolor), the neighborhood is quite large and thus the effect
is negligible. However, for the thin orbit (\thincolor), which is closest to
the orbits of nearby open clusters (see Table~\ref{tab:real_clusters}), the
neighborhood is only a few hundred $\pc$ for low-mass clusters.}
\label{fig:Rn_mc}
\end{figure*}

@@

\item To demonstrate the previous point, we measured the local Galactic
midplane along the Solar circle in three different high-resolution, zoom-in
simulations of Milky Way mass galaxies from the FIRE collaboration. We found
that the midplane varies by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
three galaxies we considered (m12i, m12f, and m12m, respectively).

>>>>

\section{Azimuthal midplane variations at the Solar Circle: estimates from
FIRE} \label{sec:local_fire}
\subsection{Description of Simulation} \label{ssec:cosmozoom}
Here we briefly describe the cosmological-hydrodynamical zoom-in simulations
used in this work. Cosmological zoom-ins allow one to simulate a selected
region at high resolution embedded in a low-resolution cosmological background
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}, to which can be added
a hydrodynamical simulation of the baryonic component with recipes for star
formation and feedback. The FIRE collaboration has simulated a number of Milky
Way-mass galaxies using this technique as part of the {\em Latte} suite of
FIRE-2 simulations \citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. For this
work we focus on the three zoom-ins considered in \citet{2018arXiv180610564S},
which show broad agreement of many of their global properties with
observations of the Milky Way. The $\z=0$ snapshots\footnote{In this work, to
avoid confusion with the vertical height $z$, we refer to cosmological
redshift as $\z$.} of these three simulations, called \mi , \mf , and \mm\ for
shorthand, are publicly available alongside associated mock {\em Gaia} DR2
catalogues generated from them.\footnote{available at
\url{http://ananke.hub.yt}}

These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
gas particles of mass $\sim 7000 \text{--} 20,000\,\Msun$, and star particles
of mass $\sim 5000 \text{--} 7000\,
\Msun$,\footnote{In fact, the simulation \mi\ contained a gas
splitting bug which causes a small number of gas and star particles to have
larger than typical masses. These constitute only $\sim0.2\%$ of the star
particles in the $\z=0$ snapshots, and are only a factor of a few more massive
than typical star particles. Since our results are broadly consistent between
the three simulations, we ignore this minor complication.} with the lower end
coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
respectively.\footnote{This is $2.8$ times the often-quoted
Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
median softening length for cold ($T < 10^4\,\text{K}$) gas particles around
roughly solar positions (with cylindrical radii within $500\,\pc$ of
$8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is $12.4, 12.5, 10.3\,\pc$ for \mi , \mf ,
and \mm , respectively.

The softening lengths used in the simulations can affect the ability to
resolve the very thinnest planar structures, which in turn can affect how much
the density-based midplane varies as a function of azimuth. The Milky Way's
dense, star-forming gas disk is thought to have a scale height of about 40 pc (3--4 times the
cold gas softening length; \citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of about 300 pc (30 times the stellar softening length; \citealt{2008ApJ...673..864J}). We therefore expect that
resolution effects may still be affecting the scale heights of these
components in the simulations, and indeed the gas and stellar scale heights of
the simulated galaxies are equal to or larger than the MW's. However, in this
work we are considering azimuthal variations in the midplane \emph{as defined
by the mass density}, which will subsequently be lower in the simulations than
in the MW: within 200 pc of the Sun the measured gas and stellar volume
densities in the MW are roughly comparable at $\sim .04\ \Msun\
\unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
Table 3 of \citealt{2018arXiv180610564S}). We thus expect that the degree of
midplane variation we find in the simulated systems is likely to be
\emph{less} than in the real MW, since on the smallest scales the simulations
are \emph{smoother} than the real Galaxy.

Cosmological simulations of Milky Way-mass galaxies are not perfect
representations of the true Milky Way in other ways as well, as discussed in
\citet{2018arXiv180610564S}. For instance, the velocity structure of \mi\ is
closer to M31's than the Milky Way's (Loebman et al. in prep). However, in
this work we are most interested in the global properties of the potential,
and specifically in deviations from axisymmetry. From this perspective, the
simulated galaxies are actually \emph{more} axisymmetric than we might expect
of the Milky Way. While they have prominent spiral arms, none has as strong a
bar as the MW does at present day, and none has a nearby companion like the
Large Magellanic Cloud. One of the three we consider (\mf) does have an
ongoing interaction with a satellite galaxy similar to Sagittarius, which has
punched through the galactic disk outside the solar circle and induced some
warping. However, we expect that, as with the use of finite smoothing lengths,
the degree of midplane variation in the simulations is likely a factor of a
few lower than what could be expected in the MW.

@@

To allow for potential small inaccuracies in the determination of the original fiducial coordinate system,
we also subtract the best fit $\text{A} \sin{\left(\phi + \text{B}\right)} +
\text{C}$ curve from the midplane as a function of azimuth. For $\text{A}$ the
values are $-174$, $62$, and $3.8\,\pc$, for $\text{B}$ the values are $0.3$,
$0.65$, and $0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$,
and $-15\,\pc$ for \mi , \mf , and \mm , respectively. For the assumed solar
radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
from the values of $\text{A}$ --- we compute $1.22$, $0.43$, and $0.03\,\deg$
for \mi , \mf , and \mm . These angle offsets are consistent with the values
given in \citet{2018arXiv180610564S} for the difference between the $z$-axis
as defined by the gas and stars.

Figure~\ref{fig:midplane} shows the inferred midplane our imaginary observer
would make as a function of azimuth for each galaxy. The 5--95 percentile range from the bootstrap procedure is shown as the dashed-line error bars. The $90\%$ interquartile
range for each galaxy is $185$, $162$, $84\,\pc$ for \mi , \mf , and \mm .

\begin{figure*}
\plotone{fig/midplane_fit.pdf}
\caption{The local midplane determined at the fiducial Solar radius
($8.2\,\kpc$) for the three FIRE galaxies \mi , \mf , and \mm\ ({\em left},
{\em center}, and {\em right} panels). The local midplane is determined at a
position $\phi$ by taking the median height of all stars within $R=0.5\,\kpc$
and $z=1\,\kpc$ (in cylindrical coordinates). The procedure is performed again
using the new height $10$ times to converge on the local midplane height. In
order to allow for the possibility that the fiducial Galactocentric coordinate
system is incorrect, we subtract the best fit $A\sin{(\phi+B)}+C$ curve from
each panel --- this figure is reproduced with the original midplane
determination (i.e. before subtracting the best fit sine curve) in
Appendix~\Gus{x}. We then bootstrap resample $1000$ times on all stars within
a $2\,\kpc$ height of the fiducial midplane to determine error bars (95\uth
and 5\uth percentiles), which we report as dashed lines.}
\label{fig:midplane}
\end{figure*}

@@

\subsection{Applicability to the Milky Way} \label{ssec:is_it_real}
As we briefly discussed in Section~\ref{ssec:cosmozoom}, cosmological zoom-in
simulations of Milky Way-mass galaxies are not perfect representations of the
Milky Way, and do not reproduce every available observation of the Milky Way.
For instance, the velocity dispersion-age relation in \mi , \mf , and \mm\ is
closer to that of M31 \citep{2018arXiv180610564S}.

In order to assess the applicability of our work to the real Milky Way, we
consider a subset of Table~3 from \citet{2018arXiv180610564S} --- our
Table~\ref{tab:scale_height} shows the stellar scale heights of \mi , \mf ,
and \mm\ and the current best estimates for the Milky Way. One noticeable
feature is that the FIRE galaxies are more ``puffy'' than the Milky Way.
\Gus{Maybe FIRE collaboration has some insight into the mechanism for this?}.
This would seem to suggest that the midplane estimates from
Figure~\ref{fig:midplane} over-estimate the true variation in the Milky Way.

\begin{deluxetable}{ccc}
\tablecaption{Stellar disk scale heights of the real Milky Way and the
simulated galaxies considered in this work.\label{tab:scale_height}}
\tablehead{\colhead{galaxy} & \colhead{thin disk} & \colhead{thick disk} \\ 
\colhead{} & \colhead{(pc)} & \colhead{(pc)} } 
\startdata
MW\tablenotemark{a} & 300 & 900 \\
\mi\tablenotemark{b} & 480 & 2000 \\
\mf\tablenotemark{b} & 440 & 1280 \\
\mm\tablenotemark{b} & 290 & 1030 \\
\enddata
% \tablerefs{\citet{2016ARAA..54..529B,2018arXiv180610564S,2008ApJ...673..864J}}

@@

Another important aspect to consider when comparing the simulated galaxies to
the Milky Way is the environment/merger history. For instance, \mi\ and \mm\
both lack a massive neighbor analogous to the LMC. On the other hand, \mf\ has
had a recenter merger with \Gus{something}. \Gus{Expand more here - how to
incorporate merger trees?}

@@

Figure~\ref{fig:Rn_mc} is most informative for the thin orbit (\thincolor).
The neighborhood around $\sim100\Msun$ clusters can be as small as
$\sim300\,\pc$ but up to $\sim1.2\,\kpc$ for \mi\ and \mf , though the
situation is marginally better for \mm\ with values ranging from
$\sim0.8\,\kpc$ to $\sim3\,\kpc$. The neighborhood values steadily increase
with $m_c$, topping out at $\sim5\,\kpc$ for $\sim10^4\,\Msun$ clusters in
\mi\ and \mf . In \mm , the neighborhood value reaches the maximum allowed of
$16.4\,\kpc$ for clusters of mass $\sim800\,\Msun$. The interpretation of this
is that the expected action space dispersion is so high for such high mass
clusters that the measured $z$~offset is not large enough throughout the
entire galaxy to be important.

For the thick-disk orbit (brown), the neighborhood is at the maximum value for
nearly all masses, except at the very low-mass end in \mi .

We repeat this entire procedure for each of the orbits of the real open
clusters in Table~\ref{tab:real_clusters} using the midplane variations from
\mi . We use \mi\ as a conservative choice, since it provides the greatest
midplane variation (Figure~\ref{fig:midplane}). We report the neighborhood
values for the mass of each cluster in the neighborhood column of
Table~\ref{tab:real_clusters}. Except for alphaPer, we find that the
neigbhorhood around each cluster is typically $\gtrsim 500\,\pc$. This
analysis assumes that the midplane from \mi\ can be accurately applied to
Milky Way orbits, which is not necessarily true. However, our analysis is
meant to be more illustrative than quantitative and these neighborhood numbers
should be understood as such.

\begin{figure*}
\plotone{fig/Rn_vs_mc.pdf}
\caption{The neighborhood around an open cluster of mass $m_c$ for the thin,
and thick orbits (\thincolor and \thickcolor, respectively; see
Table~\ref{tab:orbits}) and for each of the FIRE galaxies considered here ---
\mi , \mf , and \mm\ ({\em left}, {\em center}, and {\em right} panels,
respectively). Given the $z$ offset for a cluster of mass $m_c$ computed in
Figure~\ref{fig:cluster_offset}, we convert this to a chord length using the
result from Figure~\ref{fig:range_deltaphi}. We interpret this chord length as
the ``neighborhood'' around the cluster, i.e. the distance from the cluster
one could go before the $J_z$ error induced by our midplane effect is
comparable to the intrinsic action dispersion of the cluster. When the $z$
offset is larger than the maximum range in Figure~\ref{fig:range_deltaphi}, we
report the maximum chord length ($2\times8.2\,\kpc$). We aso report the same
procedure but using the $\pm1\sigma$ lines from
Figure~\ref{fig:range_deltaphi}, which we plot here as dashed lines. For the
thick orbit (\thickcolor), the neighborhood is quite large and thus the effect
is negligible. However, for the thin orbit (\thincolor), which is closest to
the orbits of nearby open clusters (see Table~\ref{tab:real_clusters}), the
neighborhood is only a few hundred $\pc$ for low-mass clusters.}
\label{fig:Rn_mc}
\end{figure*}

@@

\item To demonstrate the previous point, we measured the local Galactic
midplane along the Solar circle in three different high-resolution, zoom-in
simulations of Milky Way mass galaxies from the FIRE collaboration. We found
that the midplane varies by $\sim185, 162, 84\,\pc$ (middle $90\%$) for the
three galaxies we considered (\mi , \mf , and \mm , respectively).
